---
title: "Hiring_EDA"
author: "Qing Wen"
date: "2021/4/5"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(lme4)
```

## Read in the data

```{r}
hire_phase1 <- read_csv("~/303 final/Black-saber-consulting/data/phase1-new-grad-applicants-2020.csv")
hire_phase2 <- read_csv("~/303 final/Black-saber-consulting/data/phase2-new-grad-applicants-2020.csv")
hire_phase3 <- read_csv("~/303 final/Black-saber-consulting/data/phase3-new-grad-applicants-2020.csv")
hire_final <- read_csv("~/303 final/Black-saber-consulting/data/final-hires-newgrad_2020.csv")
```


## Hiring-Phase I

```{r}
# add indicator variable whether the applicant passes the first phase
success <- hire_phase2$applicant_id
hire_phase1 <- hire_phase1 %>% 
  mutate(pass = ifelse(applicant_id %in% success, 1, 0))

table1 <- table(hire_phase1$gender, hire_phase1$pass)
prop.table(table1, margin = 2)
prop.table(table1, margin = 1)

hire_phase1_filtered <- hire_phase1 %>% 
  filter(cover_letter == 1) %>% 
  filter(cv ==1)
```

It seems that the phase-I selection is unbiased with gender. 

### Visualizations

```{r}
hire_phase1 <- hire_phase1 %>% 
  mutate(cover_letter = as.factor(cover_letter)) %>% 
  mutate(cv = as.factor(cv))


# proportion of successful applicants within each sex group
hire_phase1 %>% 
  ggplot(aes(x = cv, fill = as.factor(pass))) +
  geom_bar(position = "fill") +
  facet_wrap(~gender) +
  scale_fill_manual(values = c("#fa8154", "#51c6b9")) +
  ylab("Proportion") +
  theme_minimal() +
  labs(fill = "Pass", x = "CV", y = "Proportion")

table(hire_phase1$cv, hire_phase1$pass)

# proportion of successful applicants vs. gender and cover letter
hire_phase1 %>% 
  ggplot(aes(x = cover_letter, fill = as.factor(pass))) +
  geom_bar(position = "fill") +
  facet_wrap(~gender) +
  scale_fill_manual(values = c("#fa8154", "#51c6b9"))+
  ylab("Proportion") +
  theme_minimal() +
  labs(fill = "Pass", x = "Cover letter", y = "Proportion")
```

- applicants without a cv did not pass phase I
- applicants without a cover letter did not pass phase I


## Build model

```{r}
hire_phase1_filtered <- hire_phase1 %>%
  filter(cv == 1) %>%
  filter(cover_letter == 1)

hire1_mod <- glm(pass ~ gpa + extracurriculars + work_experience, family = binomial(), data = hire_phase1_filtered)

hire1_mod_wg <- glm(pass ~ gender + gpa + extracurriculars + work_experience, family = binomial(), data = hire_phase1_filtered)

summary(hire1_mod)
summary(hire1_mod_wg)

lmtest::lrtest(hire1_mod, hire1_mod_wg)
```

- gender does not seem to play a role in selection process
- we pick `hire1_mod` as our model for phase-I
- __Question: whether we want to filter out all the applicants without a cv or cover letter__

## Hiring-Phase II

```{r}
success_2 <- hire_phase3$applicant_id
hire_phase2 <- hire_phase2 %>% 
  mutate(pass = ifelse(applicant_id %in% success_2, 1, 0))

# getting some summary statistics for gender
# temp<- hire_phase2 %>% 
#   filter(pass == 1 & gender == "Man") 
# mean(temp$technical_skills)
# mean(temp$writing_skills)
# mean(temp$speaking_skills)
# 
# temp1 <- hire_phase2 %>%
#   filter(pass == 1 & gender == "Woman")
# mean(temp1$technical_skills)
# mean(temp1$writing_skills)
# mean(temp1$speaking_skills)
# 
# summary(temp)
# summary(temp1)
```

### Visualizations

```{r}
par(mfrow = c(2, 2))

hire_phase2 %>% 
  ggplot(aes(x = technical_skills, fill = as.factor(pass))) +
  geom_histogram() +
  facet_wrap(~gender) +
  theme_minimal() +
  labs(fill = "Pass", x = "Technical skills")

hire_phase2 %>% 
  ggplot(aes(x = speaking_skills, fill = as.factor(pass))) +
  geom_bar() +
  facet_wrap(~gender) +
  theme_minimal() +
  labs(fill = "Pass", x = "Speaking skills")

hire_phase2 %>% 
  ggplot(aes(x = writing_skills, fill = as.factor(pass))) +
  geom_histogram() +
  facet_wrap(~gender) +
  theme_minimal() +
  labs(fill = "Pass", x = "Writing skills")

hire_phase2 %>% 
  ggplot(aes(x = leadership_presence, fill = as.factor(pass))) +
  geom_bar() +
  facet_wrap(~gender) +
  theme_minimal() +
  labs(fill = "Pass", x = "Leadership presence")


table(temp$gender, temp$technical_skills)

table2 <- table(hire_phase2$gender, hire_phase2$pass)
prop.table(table2, margin = 2)
prop.table(table2, margin = 1)
```

- Looking at the people that passed the second phase, 68% of the successful applicants are males.
- Looking at the four skills, it seems that the algorithm is fair in selecting applicants based on the skills. 
- __Question: on average, does it seem that more males are selected than females within the same score interval?__

### Verify assumptions

- the cases are independently distributed
- the dependent variable follows a binomial distribution

### Build a model

```{r}
# build a logistic model
hire2_mod <- glm(pass~technical_skills + writing_skills + leadership_presence + speaking_skills, family = binomial(), data = hire_phase2)

hire2_mod_wg <- glm(pass~gender + technical_skills + writing_skills + leadership_presence + speaking_skills, family = binomial(), data = hire_phase2)

lmtest::lrtest(hire2_mod, hire2_mod_wg)
summary(hire2_mod_wg)
summary(hire2_mod)
```

- gender does not play a role in hiring-2, as including gender in the model does not improve the model by much, and the term is not significant

## Hiring-Phase III

```{r}
# find the proportion of male/female in the successful applicants in the final phase
success_3 <- hire_final$applicant_id
hire_phase3 <- hire_phase3 %>% 
  mutate(pass_final = ifelse(applicant_id %in% success_3, 1, 0))

hire_phase3 %>% 
  mutate(average = (interviewer_rating_1 + interviewer_rating_2) / 2)%>% 
  arrange(desc(average))

hire_2_3 <- left_join(hire_phase3, hire_phase2, by = "applicant_id") 

prop.table(table(hire_2_3$gender, hire_2_3$pass_final), margin = 2)
prop.table(table(hire_2_3$gender, hire_2_3$pass_final), margin = 1)
```

- 80% of the successful applicants are males
- next-step: whether the selected applicants are the ones with the highest interview scores?

### Visualization
```{r}
hire_2_3 %>% 
  ggplot(aes(x = interviewer_rating_1, fill = as.factor(pass_final))) +
  geom_histogram() +
  facet_wrap(~gender) +
  theme_minimal() +
  labs(fill = "Pass", x = "Interview rating 1")

hire_2_3 %>% 
  ggplot(aes(x = interviewer_rating_2, fill = as.factor(pass_final))) +
  geom_histogram() +
  facet_wrap(~gender) +
  theme_minimal() +
  labs(fill = "Pass", x = "Interview rating 2")
```


### Build model
```{r}
hire3_mod <- glm(pass_final ~ interviewer_rating_1 + interviewer_rating_2, family = binomial(), data = hire_2_3)
hire3_mod_wg <- glm(pass_final ~ gender + interviewer_rating_1 + interviewer_rating_2, family = binomial(), data = hire_2_3)
summary(hire3_mod)
summary(hire3_mod_wg)
lmtest::lrtest(hire3_mod, hire3_mod_wg)
```

- __Problem: too few observations__