---
title: "Hiring_EDA"
author: "Qing Wen"
date: "2021/4/5"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(lme4)
```

## Read in the data
```{r}
hire_phase1 <- read_csv("~/303 final/Black-saber-consulting/data/phase1-new-grad-applicants-2020.csv")
hire_phase2 <- read_csv("~/303 final/Black-saber-consulting/data/phase2-new-grad-applicants-2020.csv")
hire_phase3 <- read_csv("~/303 final/Black-saber-consulting/data/phase3-new-grad-applicants-2020.csv")
hire_final <- read_csv("~/303 final/Black-saber-consulting/data/final-hires-newgrad_2020.csv")
```



## Hiring-Phase I

```{r}
# add indicator variable whether the applicant passes the first phase
success <- hire_phase2$applicant_id
hire_phase1 <- hire_phase1 %>% 
  mutate(pass = ifelse(applicant_id %in% success, 1, 0))

table1 <- table(hire_phase1$gender, hire_phase1$pass)
prop.table(table1, margin = 2)
prop.table(table1, margin = 1)
```

It seems that the phase-I selection is unbiased with gender. 

```{r}
hire_phase1 %>% 
  filter(pass == 1) %>% 
  ggplot(aes(x = gpa, fill = gender)) +
  geom_histogram() +
  theme_minimal()
```

## Build model

```{r}
hire1_mod <- glm(pass~cover_letter + cv + gpa + extracurriculars + work_experience, family = binomial(), data = hire_phase1)

hire1_mod_wg <- glm(pass~ gender + cover_letter + cv + gpa + extracurriculars + work_experience, family = binomial(), data = hire_phase1)

summary(hire1_mod)
summary(hire1_mod_wg)

lmtest::lrtest(hire1_mod, hire1_mod_wg)
```
- gender does not seem to play a role in selection process

## Hiring-Phase II

```{r}
success_2 <- hire_phase3$applicant_id
hire_phase2 <- hire_phase2 %>% 
  mutate(pass = ifelse(applicant_id %in% success_2, 1, 0))

# getting some summary statistics for gender
temp<- hire_phase2 %>% 
  filter(pass == 1 & gender == "Man") 
mean(temp$technical_skills)
mean(temp$writing_skills)
mean(temp$speaking_skills)

temp1 <- hire_phase2 %>%
  filter(pass == 1 & gender == "Woman")
mean(temp1$technical_skills)
mean(temp1$writing_skills)
mean(temp1$speaking_skills)

summary(temp)
summary(temp1)
```

```{r}
par(mfrow = c(2, 2))
temp %>% 
  ggplot(aes(x = writing_skills, fill = gender)) +
  geom_histogram() +
  facet_wrap(~gender) +
  theme_minimal()

temp %>% 
  ggplot(aes(x = speaking_skills, fill = gender)) +
  geom_histogram() +
  facet_wrap(~gender) +
  theme_minimal()

temp %>% 
  ggplot(aes(x = technical_skills, fill = gender)) +
  geom_histogram() +
  facet_wrap(~gender) +
  theme_minimal()

temp %>% 
  ggplot(aes(x = leadership_presence, fill = gender)) +
  geom_histogram() +
  facet_wrap(~gender) +
  theme_minimal()


table(temp$gender, temp$technical_skills)

table2 <- table(hire_phase2$gender, hire_phase2$pass)
prop.table(table2, margin = 2)
prop.table(table2, margin = 1)
```

- Looking at the people that passed the second phase, 68% of the successful applicants are males.
- question: is the algorithm based only on the four skills?

### Verify assumptions

- the cases are independently distributed
- the dependent variable follows a binomial distribution


### Build a model

```{r}
# build a logistic model
hire2_mod <- glm(pass~technical_skills + writing_skills + leadership_presence + speaking_skills, family = binomial(), data = hire_phase2)
hire2_mod_wg <- glm(pass~gender + technical_skills + writing_skills + leadership_presence + speaking_skills, family = binomial(), data = hire_phase2)

lmtest::lrtest(hire2_mod, hire2_mod_wg)

summary(hire2_mod)
```

- gender does not play a role in hiring-2, as including gender in the model does not improve the model by much, and the term is not significant

## Hiring-Phase III

```{r}
# find the proportion of male/female in the successful applicants in the final phase
success_3 <- hire_final$applicant_id
hire_phase3 <- hire_phase3 %>% 
  mutate(pass_final = ifelse(applicant_id %in% success_3, 1, 0))



hire_2_3 <- left_join(hire_phase3, hire_phase2, by = "applicant_id") 

prop.table(table(hire_2_3$gender, hire_2_3$pass_final), margin = 2)
prop.table(table(hire_2_3$gender, hire_2_3$pass_final), margin = 1)
```

- 80% of the successful applicants are males
- next-step: whether the selected applicants are the ones with the highest interview scores?

### Build model
```{r}
hire3_mod <- glm(pass_final~interviewer_rating_1 + interviewer_rating_2, family = binomial(), data = hire_phase3)
hire3_mod_wg <- glm(pass_final~gender + interviewer_rating_1 + interviewer_rating_2, family = binomial(), data = hire_2_3)
summary(hire3_mod)
summary(hire3_mod_wg)
lmtest::lrtest(hire3_mod, hire3_mod_wg)
```

