---
title: "exploratory analysis"
author: "Qing Wen"
date: "2021/3/27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(lme4)
```

```{r}
# read in the datasets
curr_employee <- read_csv("data/black-saber-current-employees.csv")
```

## Current employee data

```{r}
# salary distribution filled by gender
curr_1 <- curr_employee %>% 
  mutate(salary = str_replace(salary, "\\$", "")) %>% 
  mutate(salary = as.integer(str_replace(salary, ",", ""))) 
curr_1%>% 
  group_by(employee_id, gender) %>% 
  summarize(mean_wage = mean(salary)) %>% 
  ggplot(aes(x = mean_wage, fill = gender)) +
  geom_histogram() +
  theme_minimal() +
  labs(x="Mean salary", y="Number of employees", title="Distribution of mean salary of each current employees",subtitle="Classifed by genders") +
  scale_fill_manual(values = c("#FFC300", "#D5F2DF", "#538708"))

# summary statistics for different gender groups
summary(curr_1[curr_1$gender == "Man",])
summary(curr_1[curr_1$gender == "Woman",])
summary(curr_1[curr_1$gender == "Prefer not to say",])
```

```{r}
# productivity vs. salary by role_seniority
curr_1 %>% 
  ggplot(aes(x = productivity, y = salary, color = gender)) +
  geom_point() +
  theme_minimal() +
  facet_wrap(~role_seniority)

# gender vs. leadership_for_level
table(curr_1$gender, curr_1$leadership_for_level)

curr_1 %>% 
  ggplot(aes(x = productivity, y = salary, color = gender)) +
  geom_point() +
  theme_minimal() +
  facet_wrap(~leadership_for_level)
```

```{r}
# man vs woman employees current level satistication
curr_1%>% 
  group_by(employee_id, gender, leadership_for_level) %>% 
  summarize(mean_wage = mean(salary)) %>% 
  filter(gender != "Prefer not to say") %>%
  ggplot(aes(x = mean_wage, fill=leadership_for_level)) +
  facet_wrap(~gender) +
  geom_histogram() +
  theme_minimal() +
  #coord_flip() +
  stat_bin(bins=25) +
  labs(x="Mean salary", y="Number of employees", title="Distribution of mean salary of each current employees",subtitle="Classifed by satisfication", fill = "Satisfication") +
  scale_fill_manual(values = c("#FFC300", "#D5F2DF", "#538708"))
```

```{r}
# model without productivity
mod_noprod <- lmer(salary ~ leadership_for_level + (1|employee_id), data = curr_1)
# model with productivity
mod_nogender <- lmer(salary ~ productivity + leadership_for_level + (1|employee_id), data = curr_1)
lmtest::lrtest(mod_noprod, mod_nogender)

# model with gender
salary_mod <- lmer(salary ~ productivity + leadership_for_level + gender + (1|employee_id), data = curr_1)
lmtest::lrtest(mod_nogender, salary_mod)

# model with an additional predictor: role_seniority
mod_senior <- lmer(salary ~ productivity + leadership_for_level + gender + role_seniority 
                    + (1|employee_id), data = curr_1)
lmtest::lrtest(salary_mod, mod_senior)

summary(salary_mod)
fixef(salary_mod)
```

- we try to fit a linear mixed effect model with `salary` as the response, `productivity` and `leadership_for_level` as predictors. We treat `employee_id` as our random effects. A more complex model `salary_mod` is fit with the above covariates with an additional `gender` predictor. With a LRT, the result shows that the variable `gender` is significant in determining an employee's salary
- we need `gender`, `productivity`, `role_seniority` in the model


